Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-09 paper/paper.tex
Merge Christine's changes

In this paper, we report experimental results on a larger set of 8 algorithms, including new
variants, and a larger benchmark. We show that 5 of these algorithms are single best algorithms,
depending on the CPU time limit considered, and that combining preprocessing with an algorithm
selection on an instance by instance basis allows us to achieve better overall performance than any
single algorithm.

\section{Background}

\subsection{Subgraph Isomorphism}

\todo[inline]{Add some background on subgraph isomorphism.}

\subsection{Per-Instance Algorithm Selection}

@@

Battiti and Mascia~\cite{battiti_algorithm_2007} were the first to propose
algorithm portfolios for subgraph isomorphism problems. However, they only
consider a pure parallel portfolio without a selection mechanism consisting of
only two solvers. In addition, our instance set is much larger and considers
both hard and unsatisfiable instances.

\section{Experimental Setup}

\subsection{Problem Instances}

We consider 5725 instances, which are available in a simple text
format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12
classes.

\begin{itemize}
\item Class 1 contains randomly generated scale-free graphs \cite{}.
\item Classes 2 and 3 contain various kinds of graphs coming from \cite{}: class 2 contains small
    instances generated from the 50 first graphs of the benchmark, whereas class 3 contains larger
    instances where pattern graphs are coming from class 2, and target graphs are coming from the
    next 50 graphs of the benchmark.
\item Classes 4 to 8 contain randomly generated graphs coming from \cite{GraphDatabase}:
    bounded-degree graphs for classes 3 and 4, regular meshes for classes 5 and 6, and random graphs
    with uniform edge probabilities for class 7.
\item Classes 9 and 10 contain instances coming from images \cite{}.
\item Class 11 contains instances coming from meshes modeling 3D objects \cite{}.
\item Class 12 contains random graph instances belonging to the phase transition.
\end{itemize}

\subsection{Subgraph Isomorphism Algorithms}

We consider the following exact subgraph isomorphism
algorithms: VF2 \cite{Cordella:2004}, which is a simple backtracking search algorithm, especially
fast on trivially satisfiable instances; LAD \cite{Solnon:2010}, which does very strong reasoning
based on local all different redundant constraints; and the Glasgow algorithm \cite{McCreesh:2015},
which does an expensive preprocessing pass once, followed by weaker reasoning during search.

We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD
preprocessing) which is applied once, without performing a backtracking search, and detects
inconsistencies on many instances: For each pattern node $u$, we check that there exists at least
one target node $v$ such that for each neighbour $u'$ of $u$ there exists a
different neighbour $v'$
of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. LAD
preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: when
it does not detect inconsistency, the instance may either be consistent or not. The time complexity
of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target
nodes, respectively, and $p$ is the number of target edges.

The second variant of LAD is called new LAD and it is stronger than the default LAD introduced in
\cite{Solnon:2010}. It combines LAD filtering with new invariant properties borrowed from SND
\cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2
between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$,
and to add the constraint that the label of a pattern node (resp.\ edge) must be smaller than or
equal to the label of its associated target node (resp.\ edge).

The Glasgow algorithm has a parameter, which controls the lengths of paths used
when reasoning about non-adjacent vertices.  In experiments reported in
\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable
compromise -- longer paths lead to prohibitively expensive preprocessing on
larger, denser instances. This is often not the best choice on an instance by
instance basis: sometimes path-based reasoning gives no benefit at all,
sometimes considering only paths of length 2 suffices, occasionally paths of
length 4 are helpful, and even looking at paths of length 3 is relatively
expensive on some graphs.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/ecdf-glasgow}
\caption{Empirical cumulative distribution function for the Glasgow algorithm
with path lengths 1, 2, 3, and 4.}
\label{fig:ecdf-glasgow}
\end{figure}

Figure~\ref{fig:ecdf-glasgow} shows the cumulative distribution function for the
Glasgow algorithm with four different path lengths. As the instances become
harder to solve, longer paths become better. This is what we expect, as more
reasoning is expensive, but potentially increases the reduction of the search
space. In particular, the relative performance of the algorithms is monotone in
the time it takes to solve the instance -- path length 1 is best at first, but
gets worse as the time to solve the instance increases until path length 2 is
better. After that point, length 1 is never better than length 2 again. The same
is true for the other path lengths.

The figure illustrates the potential for portfolios and algorithm selection we
have. There is clearly no path length that dominates throughout. Therefore, we
consider 4 different instances of the Glasgow algorithm here, with path lengths
varying from 1 to 4.

\subsection{Features}

For both the pattern and the target graph, we consider the number of vertices,
edges and loops, the mean and maximum degree, the density, whether or not every
vertex has same degree, whether or not the graph is connected, the mean and
maximum distances between all pairs of vertices, and the proportion of vertex
pairs which are distance at least 2, 3 and 4 apart. We also consider the number
of variable-value assignments removed by LAD preprocessing in absolute terms, as
a percentage, and the minimum and maximum on a per-variable basis. This gives
us a total of 30 features.

In addition to computing feature values, the LAD preprocessing is able to detect
whether an instance is unsatisfiable. In this case, it is not necessary to run
an algorithm on the instance.

\subsection{Experimental Environment}

We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
timeout of $10^8$ milliseconds (a little over a day).

\section{Algorithm Selection Approach}

We apply two static steps to each instance before solving it. First, the
computation of the LAD features can detect unsatisfiability of an instance. If
an instance is identified as unsatisfiable, we do not process it further.
Second, we run the VF2 solver for 0.05 seconds. This solves easy instances
without the overhead of running algorithm selection and avoids potentially
making incorrect solver choices.

Computing the LAD features detects unsatisfiability on 1919 instances. Running
VF2 for 0.05 seconds removes a further 1470. We are left with 2336 hard
instances that we consider for algorithm selection.

@@

\section{Experimental Results}

Table~\ref{tab:res} shows the performance of our algorithm selection approach, compared to two
baselines. The virtual best solver is the oracle predictor that, for each instance, chooses the best
solver from our portfolio. This is the upper bound of what an algorithm selection approach can
achieve. The single best solver is the one solver from the portfolio that has the overall best
performance across the entire set of instances. We consider it a lower bound on performance.

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
\caption{Empirical cumulative distribution function of solvers, virtual best
solvers, and the LLAMA portfolio.}
\label{fig:portfolio-ecdf}
\end{figure}

\begin{table}[t]
\begin{center}
\begin{tabular}{|c||r||r|r|r||r|r|r|r||r|r|}
\hline
$t$ & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c||}{Glasgow}& VBS & $\delta$\\
&&pre&def&new&1&2&3&4&&\\\hline
$10^0$ & 300 &  \cellcolor{blue!25}1126 & 661 & 599 & 386 & 95 & 39 & 12 & 1132 & 6\\\hline
$10^1$ & 1699 & 1858 & 2101 &  \cellcolor{blue!25}2287 & 2061 & 1215 & 455 & 208 & 3388 & 101\\\hline
$10^2$ & 2834 & 1912 & 3376 & 3708 &  \cellcolor{blue!25}3943 & 3344 & 2295 & 1329 & 4636 & 693\\\hline
$10^3$ & 3458 & 1918 & 4233 & 4531 &  \cellcolor{blue!25}4908 & 4763 & 4262 & 3426 & 5170 & 262\\\hline
$10^4$ & 3701 & 1919 & 4888 & 5025 & 5156 &  \cellcolor{blue!25}5253 & 5022 & 4408 & 5332 & 79\\\hline
$10^5$ & 3850 & 1919 & 5108 & 5193 & 5301 &  \cellcolor{blue!25}5377 & 5292 & 5082 & 5434 & 57\\\hline
$10^6$ & 3977 & 1919 & 5244 & 5307 & 5386 &  \cellcolor{blue!25}5452 & 5451 & 5235 & 5500 & 48\\\hline
$10^7$ & 4082 & 1919 & 5336 & 5414 & 5458 & 5517 &  \cellcolor{blue!25}5518 & 5425 & 5567 & 49\\\hline
$10^8$ & 4191 & 1919 & 5423 & 5479 & 5508 &  \cellcolor{blue!25}5561 & 5560 & 5554 & 5608 & 47\\\hline
\end{tabular}
\end{center}
\caption{Number of solved instances at different CPU time limits: Each line displays a time limit
$t$ (in milliseconds) followed by the number of instances solved within this time limit by VF2, LAD
preprocessing, default LAD, and new LAD, Glasgow with  the lengths of paths limited to 1, 2, 3 and
4, and the VBS. The best single algorithm is highlighted in blue, and the difference between VBS and
single best if displayed in the last column ($\delta$).\label{expTime}}
\end{table}

Table \ref{expTime} displays the number of instances solved at different CPU time limits, ranging
from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time limit
considered. Simple approaches like LAD preprocessing and VF2 are able to solve easy instances very
quickly, in a few milliseconds. However, they are not able to solve harder instances: LAD
preprocessing is an incomplete approach which can only detect rather trivial unconsistencies; VF2
performs a basic backtracking search and is very fast on easy instances, because it does not compute
expensive invariants. Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2
and LAD preprocessing when considering longer time limits: 10 ms for default and new LAD and Glasgow
1, and 100 (resp. 1000 and 10000) ms for Glasgow 2 (resp. 3 and 4). The Virtual Best Solver (VBS),
which considers the best algorithm for each instance separately, obtains much better results,
showing us that the algorithms have complementary performance. The difference between VBS and single
best (column $\delta$ of Table \ref{expTime}) is more particularly important for rather small CPU
time limits. In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the
limit is 100 ms. In many applications, it is important to have the fastest possible algorithm. For
example, in pattern recognition applications, we often have to solve subgraph isomorphism problems
for a very large number of graphs (in order to find a pattern image in a large database of target
images, for example), so that having an algorithm that is able to solve an instance in 100 ms
instead of 1000 ms makes a big difference. Therefore, it is important to select the best algorithm
for each instance.

Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class
instances are coming from, unlike SAT. For all classes, there are always at least 2 algorithms which
are the best for at least one instance of the class. In particular, for classes 2 and 3, each
algorithm is the best for at least one instance (except Glasgow 4 for Class 2).

\begin{table}
\begin{center}
\begin{tabular}{|c||r||r|r|r||r|r|r|r|}
\hline
Class & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c|}{Glasgow}\\
&&pre&def&new&1&2&3&4\\\hline
1\\\hline
2\\\hline
3\\\hline
4\\\hline
5\\\hline
6\\\hline
7\\\hline
8\\\hline
9\\\hline
10\\\hline
11\\\hline
12\\\hline

\end{tabular}
\end{center}
\caption{Number of times each algorithm is best, for each class.\label{expClass}}
\end{table}

Table 3 shows the performance of our algorithm selection approach, compared to the VBS (which is the
upper bound of what an algorithm selection approach can achieve), and the single best solver, at
different CPU time limits (the single best corresponds to results highlighted in blue in Table 1).
Comment Table 3: Mean runtime but also number of solved instances at different CPU time limits.
Logically, LLAMA may solve less instances than single best for small time limits because it spends
time to compute features and choose a solver, but it should be better for larger time limits...

\begin{table}
\begin{tabular}{|l|r|rrrrrrrrr|}
&mean & \multicolumn{9}{c|}{Number of solved instances}\\
&runtime & $10^0$ &  $10^1$ &  $10^2$ &  $10^3$ &  $10^4$ &  $10^5$ &  $10^6$ &  $10^7$ &  $10^8$\\\hline
VBS & 2375.9 & 1132 & 3388 & 4636 & 5170 & 5332 & 5434 & 5500 & 5567 & 5608\\\hline
Single best & 3177.6 & 1126 & 2287 & 3943 & 4908 & 5253 & 5377 & 5452 & 5518 & 5561\\\hline
LLAMA & 2706.3\\\hline
\end{tabular}
\caption{}
\end{table}

Add Figure 1 which compares single best at $10^8$ ms (Glasgow 2) with LLAMA ?

>>>>

In this paper, we report experimental results on a larger set of 8 algorithms, including new
variants, and a larger benchmark. We show that 4 of these algorithms are single best algorithms,
depending on the CPU time limit considered, and that combining preprocessing with an algorithm
selection on an instance by instance basis allows us to achieve better overall performance than any
single algorithm.

\section{Definitions and notations}

A \emph{graph} $G=(N,E)$ consists of a \emph{node set} $N$ and an \emph{edge set} $E \subseteq N \times N$, where an edge $(u,u')$ is a couple of nodes. The number of neighbors of a node $u$ is the degree of $u$, $d^\circ(u)=\#\{ (u,u')\in E\}$. In this paper, we implicitely consider non directed graphs, such that $(u,u')\in E\Leftrightarrow (u',u)\in E$. The extension to directed graphs is rather straightforward, and all algorithms compared in this paper can handle directed graphs as well.

A \emph{subgraph isomorphism problem} between a pattern graph $G_p=(N_p,E_p)$ and a target graph $G_t=(N_t,E_t)$ consists in deciding whether $G_p$ is isomorphic to some subgraph of $G_t$. More precisely, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a different target node to each pattern node, and that preserves pattern edges, i.e., $\forall (u,u') \in E_p, (f(u),f(u')) \in E_t$.

Note that the subgraph is not necessarily induced so that two pattern nodes that are not linked by an edge may be matched to two target nodes which are linked by an edge. This problem is also called subgraph monomorphism or subgraph matching in the literature. 
%{\em Induced subgraph isomorphism} is a particular case of subgraph isomorphism such that the subgraph isomorphism function must also preserve target edges between matched nodes, i.e., 
%\[\forall (u,u') \in N_p^2, (u,u')\in E_p \Leftrightarrow(f(u),f(u')) \in E_t\]

In the following, we assume $G_p=(N_p,E_p)$ and $G_t=(N_t,E_t)$ to be the underlying instance of subgraph isomorphism problem. 
We also define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\# E_t$, and $d_p$ and $d_t$ the maximal degrees of the graphs $G_p$ and $G_t$.

\section{Subgraph isomorphism algorithms}

Subgraph isomorphism problems may be solved by a systematic exploration of the search space composed of all possible injective matchings from $N_p$ to $N_t$: starting from an empty matching, one incrementally extends a partial matching by matching a non matched pattern node to a non matched target node until either some edges are not matched by the current matching (the search must backtrack to a previous choice point and go on with another extension) or all pattern nodes have been matched (a solution has been found). To reduce the search space, this exhaustive exploration is combined with filtering techniques that aim at removing candidate couples of non matched pattern-target nodes $(u,v)\in N_p\times N_t$. Different filterings may be considered; some are stronger than others (they remove more candidate couples), but also have higher time complexities.

\paragraph{Filtering for subgraph isomorphism.}

Simple filterings basically propagate difference constraints (which ensure that the matching is injective) and edge constraints (which ensure that the matching preserves pattern edges): each time a pattern node $u\in N_p$ is matched with a target node $v\in N_t$, one removes every candidate couple $(u',v')\in N_p\times N_t$ such that either $v'=v$ (difference constraint) or $(u,u')$ is a pattern edge whereas $(v,v')$ is not a target edge (edge constraint). This simple filtering (called {\em Forward-Checking}) is very fast to achieve: in ${\cal O}(n_p)$ for difference constraints, and in ${\cal O}(d_p\cdot n_t)$ for edge constraints. It is used, for example, in McGregor's algorithm \cite{mcgregor79} and in VF2 \cite{Cordella:2004}.

R\'egin has introduced in \cite{regin} a stronger filtering for difference constraints, which ensures that all pattern nodes can be matched with different target nodes, all together. This filtering (called {\em AllDifferent Generalized Arc Consistency}) removes more candidate couples than when each difference contraint is propagated separately which Forward-Checking. However, it is also more time consuming as it is done in ${\cal O}(n_p^2\cdot n_t^2)$.

Various filterings have been introduced for edge constraints. Ullman \cite{ullman} has introduced a filtering which basically ensures that for each pattern edge $(u,u')\in E_p$ and each candidate couple $(u,v)\in N_p\times N_t$, there exists a candidate couple $(u',v')\in N_p\times N_t$ such that $(v,v')$ is a target edge. Candidate couples $(u,v)$ that do not satisfy this property are iteratively removed until a fixed point is reached. This filtering (called {\em Arc Consistency}) removes more candidate couples than Forward-Checking, but it is also more time consuming as it is done in ${\cal O}(e_p\cdot n_t^2)$ when using the algorithm AC4  \cite{MH86}.

A stronger filtering may be obtained by propagating edge constraints in a more global way, as proposed in \cite{LV02}. The idea is to check for each candidate couple $(u,v)\in N_p\times N_t$ that the number of pattern nodes adjacent to $u$ is smaller than or equal to the number of target nodes that are both adjacent to $v$ and that may be matched with nodes adjacent to $u$. This is done in ${\cal O}(n_p^2\cdot n_t^2)$. This idea has been generalized in \cite{Solnon:2010} where, for each candidate couple $(u,v)\in N_p\times N_t$, a redundant Local AllDifferent (LAD) constraint ensures that each neighbour of $u$ may be matched with a different neighbour of $v$. This is done in ${\cal O}(n_p\cdot n_t\cdot d_p^2\cdot d_t^2)$.

\paragraph{Propagation of invariant properties.}

Some filterings exploit invariant properties, {\em i.e.}, properties associated with nodes such that nodes may be matched only if they have compatible properties. A classical property is the degree: a pattern node $u\in N_p$ may be matched with a target node $v\in N_t$ only if $d^\circ(u)\leq d^\circ(v)$. This property is usually used at the beginning of the search to reduce the set of candidate couples to $\{(u,v)\in N_p\times N_t\;|\;d^\circ(u)\leq d^\circ(v)\}$.
Other examples of invariant properties are the number of cycles of length $k$ passing through the node, and the number of cliques of size $k$ containing the node, which must be smaller for a pattern node than for its matched target node.
Invariant properties may also be associated with couples of nodes. For example, the number of paths of length $k$ between two pattern nodes is smaller than or equal to the number of paths of length $k$ between the target nodes they may be matched with.

These invariant properties are used, for example:
\begin{itemize}
\item  In \cite{battiti-mascia07}, to remove candidate couples $(u,v)\in N_p\times N_t$ such that the number of paths starting from  pattern node $u$ is greater than the number of paths starting from  target node $v$~;
\item In \cite{Audemard:2014} to generalize the LAD constraint proposed in \cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be matched with all different compatible target nodes, where compatibility is defined with respect to invariant properties~;
\item In \cite{McCreesh:2015} to filter the set of candidate couples before starting the search.
\end{itemize}
In \cite{Audemard:2014} , the length of paths is not limited, and it is iteratively incremented until no more couple is removed. In \cite{battiti-mascia07} and \cite{McCreesh:2015},  algorithms are parameterized by the maximum path length considered when counting paths: Larger values for this parameter remove more candidate couples, but are also more time consuming. Experiments reported in \cite{battiti-mascia07} show us that the best setting depends on the instance considered, and that a portfolio running several randomized 
versions in time-sharing decreases the total CPU-time needed to find a solution for feasible instances. In experiments reported in \cite{McCreesh:2015}, this parameter has been set to 3, as this setting seemed to be a reasonable compromise.

\paragraph{Algorithms selected in our portfolio.}
We have selected a set of algorithms with complementary performance:
\begin{itemize}
\item VF2 \cite{Cordella:2004}, which performs a weak filtering so that it is especially fast on trivially satisfiable instances;
\item LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques (AllDifferent Generalized Arc Consistency and LAD);
\item Glasgow \cite{McCreesh:2015}, which does an expensive preprocessing based on path length invariant properties, followed by weaker filtering (forward-checking) during search. We consider 4 different instanciations of Glasgow, with path lengths varying from 1 to 4: The longer the path, the stronger the reduction, and the more expensive to achieve.
\end{itemize}

We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called incompleteLAD) which is applied once, without performing a backtracking search, and very quickly detects inconsistencies on many instances: For each pattern node $u$, we check that there exists at least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. IncompleteLAD is an incomplete algorithm as it can only detect inconsistency of some instances: when it does not detect inconsistency, the instance may either be consistent or not. As a counterpart of this incompleteness, it is very fast: its time complexity is ${\cal O}(n_p(n_t+e_t))$.

The second variant of LAD is called pathLAD and it combines the LAD constraints introduced in \cite{Solnon:2010} with the exploitation of path length properties, as proposed in \cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2 between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$, and to add the constraint that the label of a pattern node (resp. edge) must be smaller than or equal to the label of its associated target node (resp. edge).

\section{Experimental Comparison of Algorithms of the Portfolio}\label{expComp}

\paragraph{Benchmark.} We consider 5725 instances, which are available in a simple text format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12 classes.

\begin{itemize}
\item Class 1 contains randomly generated scale-free graphs \cite{constraints10}.
\item Classes 2 and 3 contain various kinds of graphs coming from \cite{LV02}: class 2 contains small
    instances generated from the 50 first graphs, whereas class 3 contains larger
    instances where pattern graphs are coming from class 2, and target graphs are coming from the
    next 50 graphs.
\item Classes 4 to 8 contain randomly generated graphs coming from \cite{GraphDatabase}:
    bounded-degree graphs for classes 3 and 4, regular meshes for classes 5 and 6, and random graphs
    with uniform edge probabilities for class 7.
\item Classes 9 and 10 contain instances coming from segmented images \cite{pr15,cviu11}.
\item Class 11 contains instances coming from meshes modeling 3D objects \cite{cviu11}.
\item Class 12 contains random graph instances belonging to the phase transition.
\end{itemize}

\paragraph{Experimental setup.} We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
timeout of $10^8$ milliseconds (a little over a day).

\begin{figure}[t]
\includegraphics[width=\textwidth]{courbe.pdf}
\caption{Evolution of the number of solved instances with respect to CPU time, for the 8 algorithms of the portfolio, and a Virtual Best Solver (VBS).\label{expTime}}
\end{figure}

\paragraph{Results.}Figure \ref{expTime} displays the evolution of the number of instances solved with respect to CPU time. It shows us that the best solver depends on the time limit
considered. IncompleteLAD is able to solve easy inconsistent instances very quickly, in a few milliseconds. Hence, for time limits lower than 5 ms, the best solver is incompleteLAD. However, it is not able to solve harder inconsistent instances, neither can it solve consistent instances.

PathLAD and Glasgow(1) outperform incompleteLAD when considering longer time limits: PathLAD is the best solver for time limits larger than 5ms and lower than 40 ms, and Glasgow(1) is the best solver for time limits larger than 40ms and lower than 3000 ms.

Finally, for CPU time limits larger than 3000 ms, the best solver becomes Glasgow(2).
As we increase the CPU time limit, variants of Glasgow with longer paths (Glasgow(3) and Glasgow(4)) become better. This is what we expect, as more reasoning is expensive, but potentially increases the reduction of the search space. In particular, the relative performance of Glasgow variants is monotone in the time it takes to solve the instance – Glasgow(1) is best at first, but gets worse as the time to solve the instance increases until Glasgow(2) is better. After that point, Glasgow(1) is never better than Glasgow(2) again. The same is true for path lengths 3 and 4.

The figure illustrates the potential for portfolios and algorithm selection we have. There is clearly no single solver that dominates throughout. 
Furthermore, the Virtual Best Solver (VBS), which considers the best algorithm for each instance separately, obtains much better results, showing us that the algorithms have complementary performance. The difference between VBS and single best is more particularly important for rather small CPU time limits (smaller than 1000 ms). In many applications, it is important to have the fastest possible algorithm. For example, in pattern recognition applications such as those reported in \cite{pr15,cviu11}, we often have to solve subgraph isomorphism problems for a very large number of graphs (in order to find a pattern image in a large database of target images, for example), so that having an algorithm that is able to solve an instance in 100 ms instead of 1000 ms makes a big difference. Therefore, it is important to select the best algorithm for each instance, even if the instance is an easy one.

Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class instances are coming from, unlike SAT. For all classes, there are always at least 2 algorithms which are the best for at least one instance of the class. In particular, for classes 2 and 3, each algorithm is the best for at least one instance (except Glasgow 4 for Class 3).

\begin{table}[t]
\begin{center}
\begin{tabular}{|c||r||r|r|r||r|r|r|r|}
\hline
Class & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c|}{Glasgow}\\
&&incomplete&default&path&1&2&3&4\\\hline\hline
1 &        0 &       20 &        0 &        0 &       80 &        0 &        0 &        0 \\\hline
2 &      201 &       92 &      189 &      270 &      520 &      180 &       53 &       15 \\\hline
3 &      112 &     1608 &      617 &      959 &      396 &      195 &       21 &        0 \\\hline
4 &      270 &        0 &        0 &        0 &        5 &        0 &        0 &        0 \\\hline
5 &      266 &        0 &        1 &        3 &       31 &        0 &        0 &        0 \\\hline
6 &       71 &        0 &        0 &        0 &        7 &       14 &        1 &        0 \\\hline
7 &      270 &        0 &        0 &        0 &        5 &        0 &        0 &        0 \\\hline
8 &        0 &        0 &        0 &        1 &      195 &       69 &        6 &        0 \\\hline
9 &       77 &        3 &        0 &       19 &      103 &        1 &        0 &        0 \\\hline
10 &       13 &        0 &        2 &        2 &        7 &        0 &        0 &        0 \\\hline
11 &        2 &      142 &       71 &       17 &       23 &        0 &        0 &        0 \\\hline
12 &        0 &        0 &        1 &        2 &      158 &        6 &        1 &        0 \\\hline\hline
Total & 1282 & 1865 & 881 & 1273 & 1530 & 465 & 82& 15\\\hline
\end{tabular}
\end{center}
\caption{Number of times each algorithm is best, for each class.\label{expClass}}
\end{table}

\section{Per-Instance Algorithm Selection for subgraph isomorphism}

@@

Battiti and Mascia~\cite{battiti-mascia07} were the first to propose
algorithm portfolios for subgraph isomorphism problems. However, they only
consider a pure parallel portfolio without a selection mechanism consisting of
only two solvers. In addition, our instance set is much larger and considers
both hard and unsatisfiable instances.

Our approach is composed of 3 steps described below: First, we run two presolvers in a static way to quickly solve easy instances; Second, we extract features from instances which are not solved by the first step; Finally, we select an algorithm and run it.

\paragraph{Static steps.}

Experimental results reported in Section \ref{expComp} have shown us that incompleteLAD is very fast (7 ms on average) and is able to very quickly solve 1919 instances. Therefore, we propose to first run incompleteLAD: If the instance is solved, then we do not process it further. Furthermore, running IncompleteLAD allows us to collect features on the number of candidate couples which are removed by the filtering.

VF2 is also able to solve very quickly many easy instances: Among the 3806 instances which are not solved by incompleteLAD, 1470 are solved by VF2 in less than 50 milliseconds. Therefore, after running incompleteLAD, we run the VF2 solver for 50 ms. This solves easy instances without the overhead of running algorithm selection and avoids potentially making incorrect solver choices. After running incompleteLAD and VF2 for 50 ms, we are left with 2336 hard instances that we consider for algorithm selection.

\paragraph{Features.}

For both the pattern and the target graph, we consider the number of vertices,
edges and loops, the mean and maximum degree, the density, whether or not every
vertex has same degree, whether or not the graph is connected, the mean and
maximum distances between all pairs of vertices, and the proportion of vertex
pairs which are distance at least 2, 3 and 4 apart. We also consider the number
of candidate couples removed by incompleteLAD in absolute terms, as
a percentage, and the minimum and maximum on a per-variable basis. This gives
us a total of 30 features.

\paragraph{Algorithm selection.}

@@

\section{Experimental Evaluation of Algorithm Selection}

Table~\ref{tab:res} shows the performance of our algorithm selection approach, compared to two
baselines. The virtual best solver is the oracle predictor that, for each instance, chooses the best
solver from our portfolio. This is the upper bound of what an algorithm selection approach can
achieve. The single best solver is the one solver from the portfolio that has the overall best
performance across the entire set of instances, at the CPU time limit of $10^8$ ms, {\em i.e.}, Glasgow(2). We consider it a lower bound on performance.

@@
